### 2024-04-19 paper/paper.qmd
Initial commit

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(modelsummary)
library(ggplot2)
library(dplyr)
library(rstanarm)
library(parameters)
library(broom)
library(broom.mixed)
```

```{r}
#| include: false
#| warning: false
#| message: false

clean_data_below40 <- read_csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/analysis_data/analysis_data_below40.csv")|>
  select(Year.x, Cause,Population,Deaths, Death_rate_per_100000, Age)

clean_data_above40 <- read_csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/analysis_data/analysis_data_above40.csv")|>
  select(Year.x, Cause,Population,Deaths, Death_rate_per_100000, Age)
```

# Introduction

# Data {#sec-data}

## Raw Data

## Clean Data

```{r}
#| label: tbl-data-below40
#| tbl-cap: Top 10 Mortality Causes in Switzerland for young adults in 2010 
#| echo: false
#| warning: false
#| message: false

#cleaned data
head(clean_data_below40, n=10)|>
  kable(
    col.names = c("Year", "Causes", "Population", "Deaths", "Death rate", "Age"), 
    align =  c("c","c","c","c","c","c")
  )
```

```{r}
#| label: tbl-data-above40
#| tbl-cap: Top 10 Mortality Causes in Switzerland for middle ages in 2010 
#| echo: false
#| warning: false
#| message: false

#cleaned data
head(clean_data_above40, n=10)|>
  kable(
    col.names = c("Year", "Causes", "Population", "Deaths", "Death rate", "Age"), 
    align =  c("c","c","c","c","c","c")
  )
```

```{r}
#| label: tbl-common_mortality_causes_below40
#| fig-cap: Common Mortality Causes for young adults in 2010,1015,and 2019
#| echo: false

top_10_2010_below40 <- clean_data_below40$Cause[clean_data_below40$Year.x == 2010]
top_10_2015_below40 <- clean_data_below40$Cause[clean_data_below40$Year.x == 2015]
top_10_2019_below40  <- clean_data_below40$Cause[clean_data_below40$Year.x == 2019]

common_causes_below40  <- intersect(intersect(top_10_2010_below40, top_10_2015_below40), top_10_2019_below40 )
# Calculate total deaths for each common cause of death
total_deaths_below40  <- aggregate(Deaths ~ Cause, subset(clean_data_below40, Cause %in% common_causes_below40), sum)

common_causes_below40_table <- total_deaths_below40  %>% 
  arrange(desc(Deaths))%>%
  head(5)

kable(common_causes_below40_table)
```
```{r}
#| label: tbl-common_mortality_causes_above40
#| fig-cap: Common Mortality Causes for middle age adults in 2010,1015,and 2019
#| echo: false

top_10_2010_above40 <- clean_data_above40$Cause[clean_data_above40$Year.x == 2010]
top_10_2015_above40 <- clean_data_above40$Cause[clean_data_above40$Year.x == 2015]
top_10_2019_above40  <- clean_data_above40$Cause[clean_data_above40$Year.x == 2019]

common_causes_above40  <- intersect(intersect(top_10_2010_above40, top_10_2015_above40), top_10_2019_above40 )
# Calculate total deaths for each common cause of death
total_deaths_above40  <- aggregate(Deaths ~ Cause, subset(clean_data_above40, Cause %in% common_causes_above40), sum)

common_causes_above40_table <- total_deaths_above40  %>% 
  arrange(desc(Deaths))%>%
  head(5)

kable(common_causes_above40_table)
```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

datasummary(
  Deaths ~ Min + Mean + Max + SD + Var + N,
  fmt = 0,
  data = clean_data_below40
)

```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

datasummary(
  Deaths ~ Min + Mean + Max + SD + Var + N,
  fmt = 0,
  data = clean_data_above40
)
```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

clean_data_below40%>%
  filter(Cause %in% common_causes_below40_table$Cause) %>%
  ggplot(aes(x = Year.x, y = Deaths, color = Cause)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Year", y = "Number of deaths")+
  scale_x_continuous(breaks = c(2010, 2015, 2019))+ 
  ylim(0,250)# Adjust the limits as needed
```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

clean_data_above40%>%
  filter(Cause %in% common_causes_above40_table$Cause) %>%
  ggplot(aes(x = Year.x, y = Deaths, color = Cause)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Year", y = "Number of deaths")+
  scale_x_continuous(breaks = c(2010, 2015, 2019))+ 
  ylim(100,650)# Adjust the limits as needed
```

# Model

## Model set-up

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

### Model justification

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

## Load models
cause_of_death_below40 <-
  readRDS(file ="~/Top-mortality-causes-of-age-groups-in-Switzerland/models/cause_of_death_below40.rds")
cause_of_death_above40 <-
  readRDS(file ="~/Top-mortality-causes-of-age-groups-in-Switzerland/models/cause_of_death_above40.rds")

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary(
  list(
    " Below 40" = cause_of_death_below40
))
```
```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary(
  list(
    " Above 40" = cause_of_death_above40
))
```
```

# Discussion

## First discussion point {#sec-first-point}

```{r}

```

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References